---
title: "Session 6: Homework 3"
author: "Your name goes here"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: flatly
    highlight: zenburn
    number_sections: yes
    toc: yes
    toc_float: yes
    code_folding: show
---


```{r, setup, echo=FALSE}
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE, 
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
options(digits = 3)

# default figure size
knitr::opts_chunk$set(
  fig.width=6.75, 
  fig.height=6.75,
  fig.align = "center"
)
```


```{r load-libraries, echo=FALSE}
library(tidyverse)  # Load ggplot2, dplyr, and all the other tidyverse packages
library(mosaic)
library(ggthemes)
library(GGally)
library(readxl)
library(here)
library(skimr)
library(janitor)
library(broom)
library(tidyquant)
library(infer)
library(openintro)
```


# Youth Risk Behavior Surveillance

Every two years, the Centers for Disease Control and Prevention conduct the [Youth Risk Behavior Surveillance System (YRBSS)](https://www.cdc.gov/healthyyouth/data/yrbs/index.htm) survey, where it takes data from high schoolers (9th through 12th grade), to analyze health patterns. In this task We will be analysing a selected group of variables from a random sample of observations during one of the years the YRBSS was conducted. We will begin by examining the correlation between physical activity and weight of the students and proceed by performing a hypothesis test to understand whether the obtained difference is indeed statistically significant.

## Load the data

We begin by loading the data. This data is part of the `openintro` textbook and we can load and inspect it. There are observations on 13 different variables, some categorical and some numerical, as shown below.

```{r}
data(yrbss)
glimpse(yrbss)
```


## Exploratory Data Analysis

We can further skim through the dataset to get a feel for missing values, summary statistics of numerical variables, and a very rough histogram.

```{r, eda_on_weight}
skim(yrbss)
```

As it can be observed, the weights of students are distributed heavily right skewed. Of course, such an observation comes at no surprise as in fact one would expect the vast majority of students to weight at around 40 to 50kg, with thos that weigh more than 80kg being the exception. Mreover, we can see that we are missing 1004 observations.

Next, we proceed to examin the possible relationship between a high schooler’s weight and their physical activity by plotting the data to allow us to quickly visualize trends and identify strong associations.To obtain a plot we first need to create a new variable `physical_3plus`, which will be `yes` if they are physically active for at least 3 days a week, and `no` otherwise.

  
```{r}
yrbss <- yrbss %>% 
  mutate(physical_3plus = ifelse(physically_active_7d >= 3, "yes", "no"))

yrbss_summary <- yrbss %>% filter(!is.na(physical_3plus)) %>% 
  group_by(physical_3plus) %>% 
  summarise(count = n()) %>% 
  mutate(prop= count/sum(count))
yrbss_summary

```
Can you provide a 95% confidence interval for the population proportion of high schools that are *NOT* active 3 or more days per week?

```{r, boxplot}
#sum population
count_pop <- yrbss_summary  %>%  summarise(sum = sum(count))
#calculating the SE
se_pop <- sqrt(yrbss_summary$prop[1] * (1-yrbss_summary$prop[1])/count_pop)
#calculating the mean
#mean_pop <- se_pop/sqrt(count_pop)
#t-critical value
t_critical <- qt(0.975, count_pop$sum-1)
lower_ci <- yrbss_summary$prop[1] - t_critical * se_pop
upper_ci <- yrbss_summary$prop[1] + t_critical * se_pop

```

Following the above analysis, it can be concluded that at the 95% confidence level the true population proportion of high school students that are not active 3 or more days per week lies between 32.3% and 33.9%. 

To proceed in our analysis we will visually examine the correlation between the weight of students with their relative physical activity. 

Make a boxplot of `physical_3plus` vs. `weight`. Is there a relationship between these two variables? What did you expect and why?

```{r, boxplot}
#boxplot
yrbss_clean <- na.omit(yrbss)
boxplot_pop <- ggplot(data = yrbss_clean, aes(x = physical_3plus,y= weight)) +
geom_boxplot()+
theme_bw()+
labs(title = "There is a small positive correlation between the activity level of students\nand their weight",
     subtitle = "Correlation between weight and activity level",
     x = "Students with more than 3 days per week of physical activity", y = "Weight (kg)")
boxplot_pop
 
#calculating the mean and the median of the weight
yrbss_clean %>%  group_by(physical_3plus)%>% summarize(mean  = mean(weight), median = median(weight))
```

The figure above illustrates that there is indeed a positive correlation between the ones doing more exercise and weight, albeit small. Even though at first this result seems rather odd, it can be explained when considering that during the ages of 9th through 12th grade most children tend to add muscles. As a result, those more active will add up significantly more muscle and thus weight more.  Nevertheless, at that point we cannot be certain on whether this result is statistically significant.

## Confidence Interval

The Boxplots above shows how the medians of the two distributions compare, but we can also compare the means of the distributions using either a confidence interval or a hypothesis test. Lets first calculate a few statistics for each group of the population, namely the students that do more than 3 days per week a physical activity, and those that don't.


```{r}
yrbss %>%
  group_by(physical_3plus) %>%
  filter(!is.na(physical_3plus)) %>% 
  summarise(mean_weight = mean(weight, na.rm = TRUE),
            sd_weight = sd(weight, na.rm=TRUE),
            count = n(),
            se_weight = sd_weight/sqrt(count),
            t_critical = qt(0.975, count-1), 
            margin_of_error = t_critical * se_weight,
            lower = mean_weight - t_critical * se_weight,
            upper = mean_weight + t_critical * se_weight
            )

```

As it can be seen, there is an observed difference of about 1.77kg (68.44 - 66.67) between the two groups, and we notice that the two confidence intervals do not overlap. It seems that the difference is at least 95% statistically significant, but we will conduct a hypothesis test to establish our conclusion.

## Hypothesis test with formula

Let's first state the null and the alternative hypothesis:

N0: mean of the first group is equal to the mean of the second group
N1: means are not equal

```{r}
t.test(weight ~ physical_3plus, data = yrbss)
```
We can conclude that at the 95% confidence level the two means are indeed different and we can reject the null hypothesis.

## Hypothesis test with `infer`

Finally, we will perform a similar hypothesis test, but this time using the `hypothesize` function, that falls into the infer workflow. 
Let's first initialize the test and simulate the test on the null distribution to test for independence of the two samples, meaning that there is no difference between the two population means.

```{r}
obs_diff <- yrbss %>%
  specify(weight ~ physical_3plus) %>%
  calculate(stat = "diff in means", order = c("yes", "no"))

```

```{r}
null_dist <- yrbss %>%
  specify(weight ~ physical_3plus) %>%
  hypothesize(null = "independence") %>%
  generate(reps = 1000, type = "permute") %>%
  calculate(stat = "diff in means", order = c("yes", "no"))

```

We can visualize this null distribution below.

```{r}
ggplot(data = null_dist, aes(x = stat)) +
  geom_histogram()+
  theme_bw()+
labs(title = "The null distribution is closely approximating a normal distribution",
     subtitle = "Plot of the simulated null distribution",
     x = "Difference in mean weight between the two groups of students", y = "Count")
  

```


Now that the test is initialized and the null distribution formed, we can visualise to see how many of the null permutations created through the simulation have a difference of at least the observed value of obs_diff = 1.77kg and calculate the respective p-value.

```{r}

null_dist %>% visualize() +
  shade_p_value(obs_stat = obs_diff, direction = "two-sided") + 
   theme_bw()+
labs(title = "The observed difference in mean weights is significantly higher than the values of the simulated Null distribution",
     subtitle = "Plot of the simulated null distribution",
     x = "Difference in mean weight between the two groups of students", y = "Count")

null_dist %>%
  get_p_value(obs_stat = obs_diff, direction = "two_sided")


```

As it can be clearly shown the observed difference is significantly higher than the values of the simulated null distribution and we thus can reject the null hypothesis that the mean weights of the two groups is equal. We thus conclude that at the 95% significance level the mean weight of students that are active more than 3 days per week is higher than that of the students that are not so active.
